Fig2: Candidate genes

### 2020-07-31 analysis/03_datawrangling.Rmd
new candidate genes, show external

# sequential in 2 or more comparisons
wideDEGsSequential <- wideDEGscandidates %>%
  select(sextissue,gene, "control_bldg", "bldg_lay", "lay_inc.d3", "inc.d3_inc.d9",
         "inc.d9_inc.d17", "inc.d17_hatch", "hatch_n5", "n5_n9") %>%
  mutate(totalNA = rowSums(is.na(.))) %>%
  filter(totalNA < 8) %>%
  arrange(gene, sextissue) %>%
  select(-totalNA) %>% 
  replace(., is.na(.), "")
wideDEGsSequential

>>>>

# sequential in 2 or more comparisons
wideDEGsSequential <- wideDEGscandidates %>%
  select(sextissue,gene, "control_bldg", "bldg_lay", "lay_inc.d3", "inc.d3_inc.d9",
         "inc.d9_inc.d17", "inc.d17_hatch", "hatch_n5", "n5_n9") %>%
  mutate(totalNA = rowSums(is.na(.))) %>%
  filter(totalNA < 8) %>%
  arrange(totalNA, gene, sextissue) %>%
  #arrange( gene, sextissue) %>%
  select(-totalNA) %>% 
  replace(., is.na(.), "")
wideDEGsSequential

### 2020-07-31 analysis/figure-2.Rmd
new candidate genes, show external

```{r candidategenes}
parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% select(-X1)
head(parentalcaregenes)

curleychampagnegenes <- parentalcaregenes %>% distinct(literature) %>% drop_na() %>% pull(literature)
GOgenes <- parentalcaregenes %>% distinct(GO)  %>% drop_na() %>% pull(GO)
candidategenes <- parentalcaregenes %>% pull(gene) 
candidategenes
```

@@

# summary DEG results from DESeq2
candidateDEGS <- read_csv("../results/suppltable1.csv") %>%
  filter(gene %in% candidategenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", "gonad" = "G")) %>%
  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
  select(gene, res, comparison)  %>%
  group_by(gene,  comparison) %>%
  summarize(res = str_c(res, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = res) %>%
  select(gene, bldg_lay, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
candidateDEGS

## table 1 summary candidate genes
table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
  select(gene, bldg_lay:n5_n9, literature, GO, NCBI) %>%
  mutate(literature = if_else(is.na(literature), " ", "X"),
         GO = if_else(is.na(GO), " ", "X"))
head(table1)
```

@@

```{r candidatevsds}
# load `candidatevsd` with `source("../R/wrangledata.R")`
candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
  select(-X1) %>%
  filter(treatment %in% charlevels) %>%
  drop_na()
head(candidatevsd)

@@

```{r correlations}

hyp1 <- makecorrdf("female", "hypothalamus", curleychampagnegenes)  
pit1 <- makecorrdf("female", "pituitary", curleychampagnegenes)  
gon1 <- makecorrdf("female", "gonad", curleychampagnegenes)

hyp2 <- makecorrdf("male", "hypothalamus", curleychampagnegenes)  
pit2 <- makecorrdf("male", "pituitary", curleychampagnegenes)  
gon2 <- makecorrdf("male", "gonad", curleychampagnegenes)

@@

candidateboxplot <- function(whichtissue, whichgenes, whichsex){

  p <- candidatevsd %>%
    filter(tissue %in% whichtissue,
           gene %in% whichgenes,
           sex %in% whichsex) %>%
    mutate(treatment = factor(treatment, levels = charlevels)) %>%
    ggplot(aes(x = treatment, y = counts)) +
    geom_boxplot(aes(fill = treatment, color = sex), outlier.shape = NA) +
    geom_jitter(size = 0.25, aes(color = sex)) +
    #facet_wrap(~sex,  nrow = 1) +
    scale_fill_manual(values = allcolors) +
    scale_color_manual(values = allcolors) +
    theme_B3() + 
    theme(legend.position = "none",
          axis.title.y = element_text(face = "italic"),
          axis.text.x = element_blank(),
          axis.ticks = element_blank(),
          axis.line.x = element_blank()) +
    labs(y = whichgenes,
         x = NULL) +
    geom_signif(comparisons = list( c( "control", "bldg"),
                                    c( "bldg", "lay"),
                                   c( "lay", "inc.d3"),
                                   c("inc.d3", "inc.d9"),
                                   c( "inc.d9", "inc.d17"),
                                   c( "inc.d17", "hatch"),
                                   c("hatch", "n5"),
                                   c( "n5", "n9")),  
                map_signif_level=TRUE,
                textsize = 1.5, family = 'Helvetica',
                vjust = 1.5, size = 0.2)

  return(p)
}

@@

```{r writefiles}
#write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
#write.csv(candidatevsd, "../results/candidatecounts.csv")
write.csv(table1, "../results/table1.csv")

@@

pdf(file="../figures/supplfig-2.pdf", width=7.25, height=7.25)
plot(supplfig2)
dev.off()
```

>>>>

```{r candidategenes}
parentalcaregenes <- read_csv("../metadata/03_parentalcaregenes.csv") %>% select(-X1)
candidategenes <- parentalcaregenes %>% pull(gene) 
candidategenes

GOgenes <- parentalcaregenes %>% distinct(GO)  %>% drop_na() %>% pull(GO)
GOgenes

literaturegenes <- parentalcaregenes %>% 
  distinct(literature) %>% 
  drop_na() %>% 
  pull(literature)
literaturegenes

litNotGO <-  literaturegenes[!literaturegenes %in% GOgenes]
litNotGOknown <- litNotGO[!litNotGO %in% c("ADRA2A", "MEST", "SLC6A4")]
litNotGOknown
```

@@

# summary DEG results from DESeq2
candidateDEGS <- read_csv("../results/suppltable1.csv") %>%
  filter(gene %in% candidategenes) %>%
  mutate(posneg = ifelse(lfc >= 0, "+", "-"),
         sex = recode(sex, "female" = "F", "male" = "M" ),
         tissue = recode(tissue, 
                         "hypothalamus" = "H",
                         "pituitary" = "P", "gonad" = "G")) %>%
  mutate(res = paste(sex, tissue, posneg, sep = "")) %>%
  select(gene, res, comparison)  %>%
  group_by(gene,  comparison) %>%
  summarize(res = str_c(res, collapse = " ")) %>%
  pivot_wider(names_from = comparison, values_from = res) %>%
  select(gene, lay_inc.d3, inc.d3_inc.d9, inc.d9_inc.d17, hatch_n5, n5_n9)
candidateDEGS

## table 1 summary candidate genes
table1 <- left_join(candidateDEGS, parentalcaregenes) %>%
  select(gene, lay_inc.d3:n5_n9, literature, GO, NCBI) %>%
  mutate(literature = if_else(is.na(literature), " ", "X"),
         GO = if_else(is.na(GO), " ", "X"))
(table1)
```

@@

```{r candidatevsds}
# load `candidatevsd` with `source("../R/wrangledata.R")`
candidatevsd <- read_csv("../results/03_candidatevsd.csv") %>% 
  select(-X1) %>%
  filter(treatment %in% charlevels) %>%
  mutate(treatment = factor(treatment, levels = charlevels)) %>%
  drop_na()
head(candidatevsd)

@@

```{r correlations}
hyp1 <- makecorrdf("female", "hypothalamus", litNotGOknown)  
pit1 <- makecorrdf("female", "pituitary", litNotGOknown)  
gon1 <- makecorrdf("female", "gonad", litNotGOknown)

hyp2 <- makecorrdf("male", "hypothalamus", litNotGOknown)  
pit2 <- makecorrdf("male", "pituitary", litNotGOknown)  
gon2 <- makecorrdf("male", "gonad", litNotGOknown)

@@

@@

```{r newHPGfig}

a <- candidateboxplot("gonad", c("COMT"), "female") + 
  labs(title =  "Females", 
       subtitle = "Hypothlamic expression" )

b <- externalboxplots("gonad", c("COMT"), "female") + 
  labs(title =  "Females", 
       subtitle = "Hypothlamic expression" )

c <- candidateboxplot("pituitary", c("PRL"), "female") + 
  labs(subtitle = "Pituitary expression" )

d <- externalboxplots("pituitary", c("PRL"), "female") + 
   labs(subtitle = "Pituitary expression" )

e <- candidateboxplot("gonad", c("AVPR1A"), "female") + 
  labs( subtitle = "Gonad expression" ) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

f <- externalboxplots("pituitary", c("AVPR1A"), "female") + 
   labs( subtitle = "Gonad expression" ) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

females <- plot_grid(a,b,c,d,e,f, ncol = 2, rel_widths = c(2,1))

a2 <- candidateboxplot("gonad", c("COMT"), "male") + 
  labs(title =  "Males", 
       subtitle = "Hypothlamic expression" )

b2 <- externalboxplots("gonad", c("COMT"), "male") + 
  labs(title =  "Males", 
       subtitle = "Hypothlamic expression" )

c2 <- candidateboxplot("pituitary", c("PRL"), "male") + 
  labs(subtitle = "Pituitary expression" )

d2 <- externalboxplots("pituitary", c("PRL"), "male") + 
   labs(subtitle = "Pituitary expression" )

e2 <- candidateboxplot("gonad", c("AVPR1A"), "male") + 
  labs( subtitle = "Gonad expression" ) + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

f2 <- externalboxplots("pituitary", c("AVPR1A"), "male") + 
   labs( subtitle = "Gonad expression" )  + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

males <- plot_grid(a2,b2,c2,d2,e2,f2, ncol = 2, rel_widths = c(2,1))

newfig2 <- plot_grid(females, males)
newfig2
```

```{r writefiles}
#write.csv(candidatevsd, "../../musicalgenes/data/candidatecounts.csv")
#write.csv(candidatevsd, "../results/candidatecounts.csv")
write.csv(table1, "../results/table1-v2.csv")

@@

pdf(file="../figures/supplfig-2.pdf", width=7.25, height=7.25)
plot(supplfig2)
dev.off()

pdf(file="../figures/fig2-2.pdf", width=7.25, height=7.25)
plot(newfig2)
dev.off()
```
